---
title: "Step 4: Renovate codebooks"
subtitle: "Now we have a profile of they the metadata linkage for a dataset we can start preparing those individual linkage specific codebooks."
pagetitle: "Renovation - Step 4"
execute: 
  echo: false
  warning: false
  error: false
editor_options: 
  chunk_output_type: console
---
 
```{r}
## Data prep
library(tidyverse)
library(reactable)
library(shiny)

load("../fair/files/quarto-f2-objects.rdata")
load("../fair/files/quarto-f3-objects.rdata")
load("../fair/files/renovation-objects.rdata")

## Reactable theming
salurbal_reactable = reactableTheme(
  color = "#5e666d",
  backgroundColor = "#e6ebee",
  borderColor = "black",
  stripedColor = "black",
  highlightColor = "bfc5c9",
  headerStyle = list(
    fontWeight = "bolder",
    fontSize = 'larger',
    borderColor = "#555"
  ),
  searchInputStyle = list(width = "98%",
                          backgroundColor = "#e6ebee",
                          margin = "auto",
                          marginTop = "3px",
                          marginBottom = "3px")
)

## Row styling 
get_row_color = function(description_tmp){
  colorTmp = case_when(
    str_detect(description_tmp,"IDENTIFIERS")~"#f4d7ad",
    str_detect(description_tmp,"STRATA DETAILS")~"#b3e9d5",
    str_detect(description_tmp,"CATEGORIZATION")~"#a3ccdf",
    str_detect(description_tmp,"INTERNAL")~"#ebd7f3",
    str_detect(description_tmp,"DETAILS")~"#c5fbc5",
    TRUE~"#e6ebee"
  )
}
## Local objects
strata_vars = c('var_name','strata_1_name','strata_1_value','strata_2_name','strata_2_value')
```


:::{.callout-important}

## Things you will need 

- **Step 4 template**:  <a href="../fair/files/codebook_template.csv" download>📥 codebook_template.csv</a>
- **[Codebook field definitions](../fair/findable-f2.qmd#codebook-codebook)**
:::

# Description


After we have evaluated the metadata linkage for a dataset. We will know which codebook and codebook variations to prepare. For each dataset we could potentiall have up to four:

1. **codebook_simple.csv**: (Very common) will link to the data via only a single identifer  `var_name` and contain all the metadata fields that were categorized as 'simple'.
2. **codebook_by_country.csv** (Very common) will link to the data via  `var_name` and `iso2`; it will contain all the metadata fields that were categorized as 'by_country'

3. **codebook_by_year.csv** (Uncommon?) will link to the data via  `var_name` and `year`; it will contain all the metadata fields that were categorized as 'by_country'
4. **codebook_by_country.csv** (Rare) will link to the data via  `var_name` and `strata_id`; it will contain all the metadata fields that were categorized as 'by_country'.
Metadata links to data via


## Deliverable


**For each data set review the metadta linkage evaluation. For each unique linkage that is present in your dataset you will need to prepare the assosiated codebook.** The thought process and deliverables for our three example datasets can be seen below. Note these codebooks are from an older template, please use the template provided below. 

:::{.callout-important}
Step 4 template: <a href="../fair/files/codebook_template.csv" download>📥 codebook_template.csv</a>
:::





:::: {.column-screen-inset-right}

::: {.panel-tabset}


##  Air Pollution (APS)

**Based on the <a href="../fair/files/ex_APS_codebook_evaluation.csv" download>APS codebook evaluation</a> we saw that all the metadata are categorized as simple; therefor our step -4-codebooks deliverable for this dataset contains one file - codebook_simple.csv**


:::::  {.panel-tabset}

#### codebook_simple.csv


```{r}
aps_simple %>% 
  reactable(
    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```

:::::  


## SEC Census dataset (CNS) 



**Based on the <a href="../fair/files/ex_CNS_codebook_evaluation.csv" download>CNS codebook evaluation</a> we saw:**

- 17 simple fields
- 2 by_country fields (source, public)

**Therefore we need to prepare two codebooks for this dataset (see below).**


:::::  {.panel-tabset}

#### codebook_simple.csv 


```{r}
cns_simple %>%
  reactable(
    pagination = FALSE,
    highlight = TRUE,
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```

#### codebook_by_iso2.csv


```{r}
cns_iso2 %>% 
  reactable(
    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```
:::::  

## Health Survey (SVY) 


**Based on the <a href="../fair/files/ex_SVY_codebook_evaluation.csv" download>SVY codebook evaluation</a> we saw:**

- 15 simple fields
- 2 by_country fields (source, public, censor)
- 2 by_strata (var_def, interpretation)

**Therefore we need to prepare three codebooks for this dataset (see below).**



:::::  {.panel-tabset}

#### codebook_simple.csv 


```{r}
svy_simple %>%
  reactable(
    pagination = FALSE,
    highlight = TRUE,
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```

#### codebook_by_iso2.csv


```{r}
svy_iso2 %>% 
  reactable(
    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```


#### codebook_by_strata.csv


```{r}
svy_strata %>% 
  reactable(
    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(minWidth = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      description = colDef(
        minWidth = 350
      )
    ),
    theme = salurbal_reactable
  )
```
:::::  

:::

::::


