---
title: "Step 2: summarize strata information"
subtitle: "If your data is stratified you will need prepare identifiers for each strata level"
pagetitle: "Renovation - Step 2"
execute: 
  echo: false
  warning: false
  error: false
editor_options: 
  chunk_output_type: console
---

```{r}
## Data prep
library(tidyverse)
library(reactable)
library(shiny)
library(htmltools)

load("../fair/files/quarto-f2-objects.rdata")
load("../fair/files/quarto-f3-objects.rdata")
load("../fair/files/renovation-objects.rdata")

## Reactable theming
salurbal_reactable = reactableTheme(
  color = "#5e666d",
  backgroundColor = "#e6ebee",
  borderColor = "black",
  stripedColor = "black",
  highlightColor = "bfc5c9",
  headerStyle = list(
    fontWeight = "bolder",
    fontSize = 'larger',
    borderColor = "#555"
  ),
  searchInputStyle = list(width = "98%",
                          backgroundColor = "#e6ebee",
                          margin = "auto",
                          marginTop = "3px",
                          marginBottom = "3px")
)

## Row styling 
get_row_color = function(description_tmp){
  colorTmp = case_when(
    str_detect(description_tmp,"IDENTIFIERS")~"#f4d7ad",
    str_detect(description_tmp,"STRATA DETAILS")~"#b3e9d5",
    str_detect(description_tmp,"CATEGORIZATION")~"#a3ccdf",
    str_detect(description_tmp,"INTERNAL")~"#ebd7f3",
    str_detect(description_tmp,"DETAILS")~"#c5fbc5",
    TRUE~"#e6ebee"
  )
}

## Local objects
strata_vars = c('var_name','strata_1_name','strata_1_value','strata_2_name','strata_2_value')
```



:::{.callout-important}
If your data is not stratified you can skip this step and move on to Step 3. 

Step 2 template:  <a href="../fair/files/strata.csv" download>📥 strata.csv</a>
:::


#  Deliverable: `strata.csv`

**`strata.csv` is a table that contains all possible strata_id for each variable. This will organize strata information 'long' meaning if a variable is stratified there should be multiple rows per variable. It should contain the following columns **  

- `var_name`
- `strata_1_name`: name of the first strata. Should have no spaces and no underdashes '_' all text should be in Pascal case.
- `strata_1_raw` raw value of the first strata. Should have no spaces and no underdashes '_' all text should be in Pascal case.
- `strata_1_value` recoded value  of the first strata (for interpretability). Should have no spaces and no underdashes '_' all text should be in Pascal case.
- `strata_2_raw` raw value of the second strata. Should have no spaces and no underdashes '_' all text should be in Pascal case.
- `strata_2_name`  recoded value  of the second strata (for interpretability). Should have no spaces and no underdashes '_' all text should be in Pascal case.
- `strata_2_value` value of the second strata. Should have no spaces and no underdashes '_' all text should be in Pascal case.


::: {.column-screen-inset-right}
::: {.panel-tabset}

##  Air Pollution Dataset (APS)




```{r}

htmltools::browsable(
  tagList(
    tags$button("Download strata.csv for APS", onclick = "Reactable.downloadDataCSV('step2_aps')"),
    step2_aps  %>% 
      mutate(strata_1_raw = NA,
             strata_2_raw = NA) %>%
      select(var_name, 
             strata_1_name, strata_1_raw, strata_1_value, 
             strata_2_name, strata_2_raw, strata_2_value) %>% 
      reactable(
        elementId = "step2_aps",
        style = list(fontFamily = "Work Sans, sans-serif", fontSize = "0.7rem"),
        pagination = FALSE, 
        highlight = TRUE,  
        sortable = F,
        height = 500,
        defaultPageSize = 20,
        defaultColDef = colDef(minWidth = 150),
        columns = list(
          field = colDef(
            width  = 150
          )
        ),
        theme = salurbal_reactable
      )
    
  )
)

```

## SEC Census dataset (CNS) 


```{r}

htmltools::browsable(
  tagList(
    tags$button("Download strata.csv for CNS", onclick = "Reactable.downloadDataCSV('step2_cns')"),
    
    step2_cns   %>% 
      mutate(strata_1_raw = strata_1_value %>% recode("Overall"="9",
                                                      "Male"="1",
                                                      "Female"="0",
                                                      "Ratio"= "ratio"),
             strata_2_raw = NA) %>%
      select(var_name, 
             strata_1_name, strata_1_raw, strata_1_value, 
             strata_2_name, strata_2_raw, strata_2_value) %>% 
      reactable(
        elementId = "step2_cns",
        style = list(fontFamily = "Work Sans, sans-serif", fontSize = "0.7rem"),
        pagination = FALSE, 
        highlight = TRUE,  
        sortable = F,
        height = 500,
        defaultPageSize = 20,
        defaultColDef = colDef(minWidth = 150),
        columns = list(
          field = colDef(
            width  = 150
          )
        ),
        theme = salurbal_reactable
      )
    
    
  )
)
```

## Health Survey (SVY) 

```{r}

htmltools::browsable(
  tagList(
    tags$button("Download strata.csv for SVY", onclick = "Reactable.downloadDataCSV('step2_svy')"),
    step2_svy  %>% 
      mutate(strata_1_raw = strata_1_value %>% recode("Overall"="9",
                                                      "Male"="1",
                                                      "Female"="0"),
             strata_2_raw = strata_2_value) %>%
      select(var_name, 
             strata_1_name, strata_1_raw, strata_1_value, 
             strata_2_name, strata_2_raw, strata_2_value) %>% 
      reactable(
        elementId = "step2_svy",
        style = list(fontFamily = "Work Sans, sans-serif", fontSize = "0.7rem"),
        pagination = FALSE, 
        highlight = TRUE,  
        sortable = F,
        height = 500,
        defaultPageSize = 20,
        defaultColDef = colDef(minWidth = 150),
        columns = list(
          field = colDef(
            width  = 150
          )
        ),
        theme = salurbal_reactable
      )
  ))
```

:::
:::
