---
title: "Step 3: evaluate metadata linkage"
subtitle: "The pairing between variable and metadata is not always one to one. For each variable we need to specify where each piece of metadata comes from."
pagetitle: "Renovation - Step 3"
execute: 
  echo: false
  warning: false
  error: false
editor_options: 
  chunk_output_type: console
---
 
```{r}
## Data prep
library(tidyverse)
library(reactable)
library(shiny)

load("../fair/files/quarto-f2-objects.rdata")
load("../fair/files/quarto-f3-objects.rdata")

## Reactable theming
salurbal_reactable = reactableTheme(
  color = "#5e666d",
  backgroundColor = "#e6ebee",
  borderColor = "black",
  stripedColor = "black",
  highlightColor = "bfc5c9",
  headerStyle = list(
    fontWeight = "bolder",
    fontSize = 'larger',
    borderColor = "#555"
  ),
  searchInputStyle = list(width = "98%",
                          backgroundColor = "#e6ebee",
                          margin = "auto",
                          marginTop = "3px",
                          marginBottom = "3px")
)

## Row styling 
get_row_color = function(description_tmp){
  colorTmp = case_when(
    str_detect(description_tmp,"IDENTIFIERS")~"#f4d7ad",
    str_detect(description_tmp,"STRATA DETAILS")~"#b3e9d5",
    str_detect(description_tmp,"CATEGORIZATION")~"#a3ccdf",
    str_detect(description_tmp,"INTERNAL")~"#ebd7f3",
    str_detect(description_tmp,"DETAILS")~"#c5fbc5",
    TRUE~"#e6ebee"
  )
}
## Local objects
strata_vars = c('var_name','strata_1_name','strata_1_value','strata_2_name','strata_2_value')
```

:::{.callout-important}

## Things you will need 

- **Step 3 template**:  <a href="../fair/files/linkage.csv" download>📥 linkage.csv</a>
- **[Codebook field definitions](../fair/findable-f2.qmd#codebook-codebook)**
:::

# Description

The goal is to evaluate what type of linkage works best for each metadata field. Details about this step can be found in the documentation for the [F3 principle](../fair/findable-f3.qmd). 


**`linkage.csv` is a table that describes how the linkage for each of the codebook fields. Starting with this template (<a href="../fair/files/linkage.csv" download>📥 linkage.csv</a>), for each codebook field (row) you should write a value of '1' in the column cells if any variable falls under that linkage type.**

The interactive table below represents how you should fill out for the dataset you are try to process.

::: {.column-screen-inset-right}

```{r}
read_csv("../fair/files/ex_APS_codebook_evaluation.csv") %>% 
  mutate(by_var = NA) %>% 
  reactable(
    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(width = 150),
    columns = list(
      field = colDef(
        width  = 150
      ),
      by_var = colDef( 
        sortable = FALSE,
        cell = function() checkboxInput('', '',value = T )
      ),
      by_var_iso2 = colDef( 
        sortable = FALSE,
        cell = function() checkboxInput('', '', )
      ),
      by_var_year = colDef( 
        sortable = FALSE,
        cell = function() checkboxInput('', '', )
      ),
      by_var_strata = colDef( 
        sortable = FALSE,
        cell = function() checkboxInput('', '', )
      )
    ),
   
    theme = salurbal_reactable
  )
```

:::


# Deliverable: `linkage.csv`

:::{.callout-important}
Step 3 template:  <a href="../fair/files/linkage.csv" download>📥 linkage.csv</a>
:::


::: {.panel-tabset}


##  Air Pollution Dataset (APS)

All codebook fields are linkable only by_variable for the APS dataset so we for all codebook fields we only check (fill out the cell as '1') the by_var column. 

```{r}
df1 = read_csv("../fair/files/ex_APS_codebook_evaluation.csv")

htmltools::browsable(
  tagList(
    tags$button("Download linkage.csv for APS", onclick = "Reactable.downloadDataCSV('step3_aps')"),
df1  %>% 
  reactable(
         elementId = "step3_aps",

    pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(width = 150),
    theme = salurbal_reactable
  )
))
```

## SEC Census dataset (CNS) 

Most of the codebook fields in the CNS dataset are linkable only be variable except for:

- `source` vary by `var_name`+`iso2` for some variables but other do not for other variables; so this row has both `by_var` and `by_var_iso2` filled out. 

```{r}
df2 = read_csv("../fair/files/ex_CNS_codebook_evaluation.csv")

htmltools::browsable(
  tagList(
    tags$button("Download linkage.csv for CNS", onclick = "Reactable.downloadDataCSV('step3_cns')"),

      

df2  %>% 
  reactable(
       elementId = "step3_cns",
       pagination = FALSE, 
    highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(width = 150),
    theme = salurbal_reactable
  )
))
```

## Health Survey (SVY) 

Most of the codebook fields in the SVY dataset are linkable only be variable except for:

- `var_def` vary by `var_name`+`strata` for some variables but other do not for other variables; so this row has both `by_var` and `by_var_strata` filled out. 
- `source` vary by `var_name`+`iso2` for some variables but other do not for other variables; so this row has both `by_var` and `by_var_iso2` filled out. 

```{r}
df3 = read_csv("../fair/files/ex_SVY_codebook_evaluation.csv")

htmltools::browsable(
  tagList(
    tags$button("Download linkage.csv for SVY", onclick = "Reactable.downloadDataCSV('step3_svy')"),

      df3  %>% 
  reactable(
    pagination = FALSE, 
          elementId = "step3_svy",
 highlight = TRUE,  
    sortable = F,
    height = 500,
    defaultPageSize = 20,
    defaultColDef = colDef(width = 150),
    theme = salurbal_reactable
  )
  ))
```

:::
